---
title: "EDA"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, 
                      error = TRUE, 
                      tidy.opts=list(width.cutoff=80), 
                      tidy  = TRUE,
                      eval = TRUE, collapse = TRUE, comment = NA)
```

# Dataset Inspection
```{r}
d <- read.csv("academy_final.csv")
str(d)
```
Based on the rough structure above, we have the following steps

(1) Dropping not useful variables. This include
+ release_year
+ director
+ wikipedia
+ rotten_tomato


(2) In addition, we should organize the following variables
+ genres: Counting the number of genres using the variable, and then drop it
+ rating: Organize them to be only PG13, R and G (There is a NA value)
+ language: We should inspect the type of languages of the movies by category, drop it if most of them are English
+ box-office: Observe distribution of the box-office to see if impacted significantly by the year
+ genre_: The genre variables are too sparse. We can try doing two 


```{r}
# dropping columns
d <- d[, !names(d) %in% c("director", "wikipedia", "rotten_tomato")]
names(d)
```

```{r}
# replace "genre" column with genre count for each film
d$genre <- sapply(d$genre, function(g) length(unlist(strsplit(g, ", "))))
names(d)[names(d) == "genre"] <- "genre_count"
```

```{r}
# extract base rating (e.g. "PG-13", "R", "PG") from full rating string
d$rating <- gsub(" \\(.*", "", d$rating)
table(d$rating)

# filling NA value
d[which(is.na(d$rating)),]
# Film "Drive My Car" is "not rated", we classify it as R
d[which(is.na(d$rating)),]$rating <- "R"
table(d$rating)
```

```{r}
# inspect film language distribution
table(d$language)
table(d$language, d$best_picture_winner)

library(ggplot2)

ggplot(d, aes(x = language, fill = factor(best_picture_winner))) +
  geom_bar(position = "dodge") +
  labs(x = "Language", y = "Count", fill = "Won Best Picture") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# low correlation/pattern -> drop language
d <- d[, !names(d) %in% c("language")]
```
```{r}
# normalize the box office for the Netflix film (All Quiet on the Western Front (2022))
d[d$box_office_millions == 2878, ]$box_office_millions = 2878/1000000

names(d)
plot_box_office <- function(df, box_col = "box_office_millions") {
  
  # aggregate average box office by year
  box_by_year <- aggregate(df[[box_col]] ~ df$release_year, data = df, FUN = mean, na.rm = TRUE)
  colnames(box_by_year) <- c("release_year", box_col)
  
  # trend line
  print(
    ggplot(box_by_year, aes(x = release_year, y = .data[[box_col]])) +
      geom_line() + geom_point() +
      geom_smooth(method = "lm", se = TRUE, color = "blue") +
      labs(x = "Release Year", y = "Average Box Office", title = "Box Office Over Time")
  )
  
  # overall bar chart ordered by year
  print(
    ggplot(df, aes(x = reorder(film, release_year), y = .data[[box_col]])) +
      geom_bar(stat = "identity") +
      labs(x = "Film", y = "Box Office") +
      theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 5))
  )
  
  # interactive stacked bar chart
  p <- ggplot(df, aes(x = factor(release_year), y = .data[[box_col]],
                      fill = reorder(film, .data[[box_col]]),
                      text = paste("Film:", film, "<br>Box Office:", .data[[box_col]]))) +
    geom_bar(stat = "identity") +
    labs(x = "Year", y = "Box Office") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
  
  print(ggplotly(p, tooltip = "text"))
}

# call with raw
plot_box_office(d)
```

```{r}
# normalize the box office by year~~~
library(dplyr)

d <- d %>%
  group_by(release_year) %>%
  mutate(box_office_norm = box_office_millions / sum(box_office_millions, na.rm = TRUE)) %>%
  ungroup()

# call with normalized
plot_box_office(d, box_col = "box_office_norm")
```


















