---
title: "EDA"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, 
                      error = TRUE, 
                      tidy.opts=list(width.cutoff=80), 
                      tidy  = TRUE,
                      eval = TRUE, collapse = TRUE, comment = NA)
```

# Dataset Inspection
```{r}
d <- read.csv("academy_final.csv")
str(d)
```
Based on the rough structure above, we have the following steps

(1) Dropping not useful variables. This include
+ release_year
+ director
+ wikipedia
+ rotten_tomato


(2) In addition, we should organize the following variables
+ genres: Counting the number of genres using the variable, and then drop it
+ rating: Organize them to be only PG13, R and G (There is a NA value)
+ language: We should inspect the type of languages of the movies by category, drop it if most of them are English
+ box-office: Observe distribution of the box-office to see if impacted significantly by the year
+ genre_: The genre variables are too sparse. We can try doing two 


```{r}
# dropping columns
d <- d[, !names(d) %in% c("director", "wikipedia", "rotten_tomato")]
names(d)
```

```{r}
# replace "genre" column with genre count for each film
d$genre <- sapply(d$genre, function(g) length(unlist(strsplit(g, ", "))))
names(d)[names(d) == "genre"] <- "genre_count"
```

```{r}
# extract base rating (e.g. "PG-13", "R", "PG") from full rating string
d$rating <- gsub(" \\(.*", "", d$rating)
table(d$rating)

# filling NA value
d[which(is.na(d$rating)),]
# Film "Drive My Car" is "not rated", we classify it as R
d[which(is.na(d$rating)),]$rating <- "R"
table(d$rating)
```

```{r}
# inspect film language distribution
table(d$language)
table(d$language, d$best_picture_winner)

library(ggplot2)

ggplot(d, aes(x = language, fill = factor(best_picture_winner))) +
  geom_bar(position = "dodge") +
  labs(x = "Language", y = "Count", fill = "Won Best Picture") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# low correlation/pattern -> drop language
d <- d[, !names(d) %in% c("language")]
```
```{r}
# normalize the box office for the Netflix film (All Quiet on the Western Front (2022))
d[d$box_office_millions == 2878, ]$box_office_millions = 2878/1000000

names(d)
plot_box_office <- function(df, box_col = "box_office_millions") {
  
  # aggregate average box office by year
  box_by_year <- aggregate(df[[box_col]] ~ df$release_year, data = df, FUN = mean, na.rm = TRUE)
  colnames(box_by_year) <- c("release_year", box_col)
  
  # trend line
  print(
    ggplot(box_by_year, aes(x = release_year, y = .data[[box_col]])) +
      geom_line() + geom_point() +
      geom_smooth(method = "lm", se = TRUE, color = "blue") +
      labs(x = "Release Year", y = "Average Box Office", title = "Box Office Over Time")
  )
  
  # overall bar chart ordered by year
  print(
    ggplot(df, aes(x = reorder(film, release_year), y = .data[[box_col]])) +
      geom_bar(stat = "identity") +
      labs(x = "Film", y = "Box Office") +
      theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 5))
  )
  
  # interactive stacked bar chart
  p <- ggplot(df, aes(x = factor(release_year), y = .data[[box_col]],
                      fill = reorder(film, .data[[box_col]]),
                      text = paste("Film:", film, "<br>Box Office:", .data[[box_col]]))) +
    geom_bar(stat = "identity") +
    labs(x = "Year", y = "Box Office") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
  
  print(ggplotly(p, tooltip = "text"))
}

# call with raw
plot_box_office(d)
```

```{r}
# normalize the box office by year~~~
library(dplyr)

d <- d %>%
  group_by(release_year) %>%
  mutate(box_office_norm = box_office_millions / sum(box_office_millions, na.rm = TRUE)) %>%
  ungroup()

# call with normalized
plot_box_office(d, box_col = "box_office_norm")
```

```{r}
# histograms for genres

library(tidyr)
# grab all genre indicator columns
genre_cols <- names(d)[grepl("^genre_", names(d)) & names(d) != "genre_count"]

# pivot to long format for easy plotting
genre_long <- d %>%
  select(film, best_picture_winner, all_of(genre_cols)) %>%
  pivot_longer(cols = all_of(genre_cols), names_to = "genre", values_to = "has_genre") %>%
  filter(has_genre == 1) %>%
  mutate(genre = gsub("genre_", "", genre),   # clean up prefix
         genre = gsub("_", " ", genre))        # underscores -> spaces

# simple genre frequency histogram
ggplot(genre_long, aes(x = reorder(genre, -table(genre)[genre]))) +
  geom_bar() +
  labs(x = "Genre", y = "Count", title = "Genre Distribution") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# genre histogram stacked by winner status
ggplot(genre_long, aes(x = reorder(genre, -table(genre)[genre]), 
                       fill = factor(best_picture_winner))) +
  geom_bar(position = "stack") +
  labs(x = "Genre", y = "Count", fill = "Won Best Picture", 
       title = "Genre Distribution by Winner Status") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# percentage of winner within genre
ggplot(genre_long, aes(x = reorder(genre, -table(genre)[genre]), 
                       fill = factor(best_picture_winner))) +
  geom_bar(position = "fill") +
  labs(x = "Genre", y = "Proportion", fill = "Won Best Picture", 
       title = "Win Rate by Genre") +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# calculate win rate per genre to use for ordering
win_rate <- genre_long %>%
  group_by(genre) %>%
  summarise(win_rate = mean(best_picture_winner == 1)) %>%
  arrange(desc(win_rate))

ggplot(genre_long, aes(x = reorder(genre, -win_rate$win_rate[match(genre, win_rate$genre)]), 
                       fill = factor(best_picture_winner))) +
  geom_bar(position = "fill") +
  labs(x = "Genre", y = "Proportion", fill = "Won Best Picture", 
       title = "Win Rate by Genre") +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# genre count and winning rate distribution
table(d$genre_count)

table(d[d$best_picture_winner == 1, ]$genre_count)
d %>%
  group_by(genre_count) %>%
  summarise(win_rate = mean(best_picture_winner == 1, na.rm = TRUE)) %>%
  ggplot(aes(x = factor(genre_count), y = win_rate)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Number of Genres", y = "Win Rate", title = "Best Picture Win Rate by Genre Count")


```
# Correlation Map

```{r}
library(ggplot2)
library(dplyr)
library(ggcorrplot)

# 1. Filter for numeric variables AND exclude all genres
# This keeps the 'clean' numerical/award variables
numeric_data_no_genres <- d %>% 
  select(where(is.numeric)) %>% 
  select(-starts_with("genre_"))

# 2. Compute Correlation
corr_matrix <- cor(numeric_data_no_genres, use = "complete.obs")

# 3. Plot Heatmap
# Since we have fewer variables now, we can set lab = TRUE to see the exact coefficients
ggcorrplot(corr_matrix, 
           hc.order = TRUE, 
           type = "lower",
           lab = TRUE, 
           lab_size = 3,
           title = "Oscar Variable Correlation (Excluding Genres)",
           colors = c("#6D9EC1", "white", "#E46726"))

```
# Does winning GG convert to winning Oscar Best Picture

```{r}

par(mfrow=c(1,2))

# A. Golden Globe Momentum
# Does a GG win 'lock in' an Oscar?
d %>%
  group_by(won_drama_gg) %>%
  summarise(win_rate = mean(best_picture_winner)) %>%
  ggplot(aes(x = factor(won_drama_gg), y = win_rate, fill = factor(won_drama_gg))) +
  geom_col() +
  scale_x_discrete(labels = c("Lost GG-Drama", "Won GG-Drama")) +
  labs(title = "The Golden Globe Boost", y = "Best Picture Win Probability") +
  theme_minimal()


d %>%
  group_by(won_comedy_gg) %>%
  summarise(win_rate = mean(best_picture_winner)) %>%
  ggplot(aes(x = factor(won_comedy_gg), y = win_rate, fill = factor(won_comedy_gg))) +
  geom_col() +
  scale_x_discrete(labels = c("Lost GG-Comedy", "Won GG-Comedy")) +
  labs(title = "The Golden Globe Boost", y = "Best Picture Win Probability") +
  theme_minimal()
```
```{r}
library(dplyr)
library(tidyr)

# Create a summary for Drama
drama_summary <- d %>%
  group_by(won_drama_gg) %>%
  summarise(
    Total_Films = n(),
    Oscar_Wins = sum(best_picture_winner),
    Win_Rate = mean(best_picture_winner)
  ) %>%
  mutate(Category = "Drama") %>%
  rename(Won_GG = won_drama_gg)

# Create a summary for Comedy
comedy_summary <- d %>%
  group_by(won_comedy_gg) %>%
  summarise(
    Total_Films = n(),
    Oscar_Wins = sum(best_picture_winner),
    Win_Rate = mean(best_picture_winner)
  ) %>%
  mutate(Category = "Comedy") %>%
  rename(Won_GG = won_comedy_gg)

# Combine them into one professional table
gg_conversion_table <- bind_rows(drama_summary, comedy_summary) %>%
  filter(Won_GG == 1) %>% # We only care about the films that actually won the GG
  select(Category, Total_Films, Oscar_Wins, Win_Rate) %>%
  mutate(Win_Rate = scales::percent(Win_Rate, accuracy = 0.1))

print(gg_conversion_table)

```

# Does winning for best editing film convert to winning Oscar Best Picture
```{r}
# B. The 'Best Editing' Rule
# Analyzing if the film_edit award is a prerequisite
d %>%
  group_by(award_film_edit) %>%
  summarise(win_rate = mean(best_picture_winner)) %>%
  ggplot(aes(x = factor(award_film_edit), y = win_rate)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Technical Prerequisite: Best Editing", x = "Won Best Editing Award") +
  theme_light()
```
```{r}
# B. The 'Best Editing' Rule
# Analyzing if the film_edit nominee is a prerequisite
d %>%
  group_by(nominee_film_edit) %>%
  summarise(win_rate = mean(best_picture_winner)) %>%
  ggplot(aes(x = factor(nominee_film_edit), y = win_rate)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Technical Prerequisite: Best Editing", x = "Nominated for Best Editing Award") +
  theme_light()
```
# The distribution of winners in box office and tomotometer

```{r}
# C. Tomatometer vs Box Office
ggplot(d, aes(x = tomatometer, y = box_office_norm, color = factor(best_picture_winner))) +
  geom_point(aes(size = total_academy_nominations), alpha = 0.6) +
  scale_color_manual(values = c("grey", "gold"), name = "Best Picture Winner") +
  labs(title = "Critic Scores vs. Commercial Success",
       x = "Tomatometer Score",
       y = "Box Office (Normalized-Year)") +
  theme_bw()
```
```{r}
# C. Tomatometer vs Box Office
ggplot(d, aes(x = tomatometer, y = popcornmeter, color = factor(best_picture_winner))) +
  geom_point(aes(size = total_academy_nominations), alpha = 0.6) +
  scale_color_manual(values = c("grey", "gold"), name = "Best Picture Winner") +
  labs(title = "Critic Scores Public VS Expert",
       x = "Tomatometer Score",
       y = "Popcornmeter Score") +
  theme_bw()

```
```{r}
d[d$popcornmeter<25,]

```

```{r}
# D. The 'Epic' Factor: Runtime
ggplot(d, aes(x = running_time_minutes, fill = factor(best_picture_winner))) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Film Length", x = "Minutes", fill = "Winner") +
  theme_minimal()
```


```{r}
# D. The 'Epic' Factor: words count
ggplot(d, aes(x = words, fill = factor(best_picture_winner))) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Words Count", x = "Word Count", fill = "Winner") +
  theme_minimal()
```

```{r}
# D. The 'Epic' Factor: character count
ggplot(d, aes(x = characters, fill = factor(best_picture_winner))) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Character Number", x = "Character Count", fill = "Winner") +
  theme_minimal()
```


```{r}
# D. The 'Epic' Factor: character count
ggplot(d, aes(x = total_academy_nominations, fill = factor(best_picture_winner))) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Award Nominations", x = "Total Academy Nominations", fill = "Winner") +
  theme_minimal()
```
```{r}
# D. The 'Epic' Factor: character count
ggplot(d, aes(x = total_academy_award_exclude_best_pic, fill = factor(best_picture_winner))) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Awards Exclude Best Pics", x = "Total Academy Wins Exclude Best Pics", fill = "Winner") +
  theme_minimal()
```


