---
title: "EDA"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, 
                      error = TRUE, 
                      tidy.opts=list(width.cutoff=80), 
                      tidy  = TRUE,
                      eval = TRUE, collapse = TRUE, comment = NA)


# Load all necessary library
library(ggplot2)
library(plotly)
library(dplyr)
library(tidyr)
library(scales)
library(ggcorrplot)
library(tidyverse)
library(factoextra) 
library(ggrepel)
```

# Dataset Inspection

```{r}
d <- read.csv("academy_final.csv")
str(d)
```
# Data Processing

Based on the rough structure above, we have the following steps

(1) Dropping not useful variables. This include

-   release_year
-   director
-   wikipedia
-   rotten_tomato

(2) In addition, we should organize the following variables

-   genres: Counting the number of genres using the variable, and then drop it
-   rating: Organize them to be only PG13, R and G (There is a NA value)
-   language: We should inspect the type of languages of the movies by category, drop it if most of them are English
-   box-office: Observe distribution of the box-office to see if impacted significantly by the year
-   genre\_: The genre variables are too sparse. We can try doing two

```{r}
# dropping columns
d <- d[, !names(d) %in% c("director", "wikipedia", "rotten_tomato")]
names(d)
```

## Creating genre count variables
```{r}
# replace "genre" column with genre count for each film
d$genre <- sapply(d$genre, function(g) length(unlist(strsplit(g, ", "))))
names(d)[names(d) == "genre"] <- "genre_count"
```


## Tidying up rating variables
```{r}
# extract base rating (e.g. "PG-13", "R", "PG") from full rating string
d$rating <- gsub(" \\(.*", "", d$rating)
table(d$rating)

# filling NA value
d[which(is.na(d$rating)),]

# Film "Drive My Car" is "not rated", we classify it as R
d[which(is.na(d$rating)),]$rating <- "R"
table(d$rating)
```


## Inspecting film language distribution - decide to drop the variables
```{r}
# inspect film language distribution
table(d$language)
table(d$language, d$best_picture_winner)
```

```{r}
# ensure the variable is a factor with both levels explicitly defined
d$best_picture_winner <- factor(d$best_picture_winner, levels = c("0", "1"))

# plot with position_dodge and scale_x_discrete(drop = FALSE)
ggplot(d, aes(x = language, fill = best_picture_winner)) +
  geom_bar(position = position_dodge(preserve = "single")) + 
  scale_fill_discrete(drop = FALSE) + # Prevents '1' from taking '0's spot
  labs(x = "Language", y = "Count", fill = "Won Best Picture") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# low correlation/pattern -> drop language
d <- d[, !names(d) %in% c("language")]
```


## Inspecting box office
```{r}
# normalize the box office for the Netflix film (All Quiet on the Western Front (2022))
d[d$box_office_millions == 2878, ]$box_office_millions = 2878/1000000

plot_box_office <- function(df, box_col = "box_office_millions") {
  
  # aggregate average box office by year
  box_by_year <- aggregate(df[[box_col]] ~ df$release_year, data = df, FUN = mean, na.rm = TRUE)
  colnames(box_by_year) <- c("release_year", box_col)
  
  # trend line
  print(
    ggplot(box_by_year, aes(x = release_year, y = .data[[box_col]])) +
      geom_line() + geom_point() + 
      geom_smooth(method = "lm", se = TRUE, color = "blue") + # linear regression to plot the graph
      labs(x = "Release Year", y = "Average Box Office", title = "Box Office Over Time")
  )
  
  # overall bar chart ordered by year
  print(
    ggplot(df, aes(x = reorder(film, release_year), y = .data[[box_col]])) +
      geom_bar(stat = "identity") +
      labs(x = "Film", y = "Box Office") +
      theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 5))
  )
  
  # interactive stacked bar chart
  p <- ggplot(df, aes(x = factor(release_year), y = .data[[box_col]],
                      fill = reorder(film, .data[[box_col]]),
                      text = paste("Film:", film, "<br>Box Office:", .data[[box_col]]))) +
    geom_bar(stat = "identity") +
    labs(x = "Year", y = "Box Office") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
  
  print(ggplotly(p, tooltip = "text"))
}

# call with raw
plot_box_office(d)
```


###  Normalizing box office to adjust for inflation and year-related factors

```{r}
# normalize the box office by year~~~
d <- d %>%
  group_by(release_year) %>%
  mutate(box_office_norm = box_office_millions / sum(box_office_millions, na.rm = TRUE)) %>%
  ungroup()

# call with normalized
plot_box_office(d, box_col = "box_office_norm")
```
### Distribution of box office of winning film in each year
```{r}
# 1. Filter for winners only and ensure data is sorted by year
winner_data <- d %>%
  filter(best_picture_winner == 1) %>%
  arrange(release_year)

# 2. Plot the normalized box office over time
ggplot(winner_data, aes(x = release_year, y = box_office_norm)) +
  geom_line(color = "gold3", size = 1, alpha = 0.5) +  # Connection line to show trend
  geom_point(color = "gold4", size = 3) +             # Points for each winner
  geom_text_repel(aes(label = film), size = 3) + # Label films
  scale_y_continuous(labels = percent, limits = c(0, max(winner_data$box_office_norm) * 1.2)) +
  labs(
    title = "Commercial Performance of Best Picture Winners (2000-2025)",
    subtitle = "Y-axis represents the film's share of the total box office for all nominees that year",
    x = "Year",
    y = "Normalized Box Office Share"
  ) +
  theme_minimal()
```


### The relationship between box office and the winning rate
```{r}
# 1. Calculate stats for the normalized box office
box_stats <- d %>%
  filter(!is.na(box_office_norm)) %>%
  group_by(best_picture_winner) %>%
  summarise(
    avg_box = mean(box_office_norm, na.rm = TRUE),
    n = n(),
    sd_box = sd(box_office_norm, na.rm = TRUE),
    se = sd_box / sqrt(n), # Standard Error for a continuous mean
    .groups = "drop"
  ) %>%
  mutate(
    lower = pmax(0, avg_box - 1.96 * se),
    upper = avg_box + 1.96 * se
  )

# 2. Plotting Average Box Office with Error Bars
ggplot(box_stats, aes(x = factor(best_picture_winner), y = avg_box, fill = factor(best_picture_winner))) +
  geom_col(width = 0.6, alpha = 0.8) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, color = "darkred") +
  scale_x_discrete(labels = c("0" = "Non-Winner", "1" = "Winner")) +
  scale_fill_manual(values = c("gray70", "gold3")) +
  labs(
    title = "Average Normalized Box Office by Winner Status",
    subtitle = "Normalization adjusts for annual price shocks and COVID-19 impact",
    x = "Award Outcome",
    y = "Average Normalized Box Office",
    fill = "Result"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```



### Plotting normalized box office by genres

```{r}
# histograms for genres

# grab all genre indicator columns
genre_cols <- names(d)[grepl("^genre_", names(d)) & names(d) != "genre_count"]

# pivot to long format for easy plotting
genre_long <- d %>%
  select(film, best_picture_winner, all_of(genre_cols)) %>%
  pivot_longer(cols = all_of(genre_cols), names_to = "genre", values_to = "has_genre") %>%
  filter(has_genre == 1) %>%
  mutate(genre = gsub("genre_", "", genre),   # clean up prefix
         genre = gsub("_", " ", genre))        # underscores -> spaces

# simple genre frequency histogram
ggplot(genre_long, aes(x = reorder(genre, -table(genre)[genre]))) +
  geom_bar() +
  labs(x = "Genre", y = "Count", title = "Genre Distribution") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# genre histogram stacked by winner status
ggplot(genre_long, aes(x = reorder(genre, -table(genre)[genre]), 
                       fill = factor(best_picture_winner))) +
  geom_bar(position = "stack") +
  labs(x = "Genre", y = "Count", fill = "Won Best Picture", 
       title = "Genre Distribution by Winner Status") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# percentage of winner within genre
ggplot(genre_long, aes(x = reorder(genre, -table(genre)[genre]), 
                       fill = factor(best_picture_winner))) +
  geom_bar(position = "fill") +
  labs(x = "Genre", y = "Proportion", fill = "Won Best Picture", 
       title = "Win Rate by Genre") +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# calculate win rate per genre to use for ordering
win_rate <- genre_long %>%
  group_by(genre) %>%
  summarise(win_rate = mean(best_picture_winner == 1)) %>%
  arrange(desc(win_rate))

ggplot(genre_long, aes(x = reorder(genre, -win_rate$win_rate[match(genre, win_rate$genre)]), 
                       fill = factor(best_picture_winner))) +
  geom_bar(position = "fill") +
  labs(x = "Genre", y = "Proportion", fill = "Won Best Picture", 
       title = "Win Rate by Genre") +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
# Adding error bars to the winning rate of each genre
# 1. Calculate stats per genre
genre_stats <- genre_long %>%
  group_by(genre) %>%
  summarise(
    win_rate = mean(best_picture_winner == 1, na.rm = TRUE),
    n = n(),
    # Calculate Standard Error for a proportion
    se = sqrt((win_rate * (1 - win_rate)) / n),
    .groups = "drop"
  ) %>%
  mutate(
    # 95% CI calculation
    lower = pmax(0, win_rate - 1.96 * se),
    upper = pmin(1, win_rate + 1.96 * se)
  )

# 2. Plot with Error Bars
ggplot(genre_stats, aes(x = reorder(genre, -win_rate), y = win_rate, fill = win_rate)) +
  geom_col(alpha = 0.8) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.3, color = "darkred") +
  scale_y_continuous(labels = percent) +
  scale_fill_gradient(low = "steelblue1", high = "steelblue4") +
  labs(
    title = "Best Picture Win Rate by Genre",
    subtitle = "Red bars indicate 95% Confidence Intervals",
    x = "Genre",
    y = "Win Rate (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )


```


## Inspecting genre count variable
```{r}
# genre count and winning rate distribution
table(d$genre_count)

table(d[d$best_picture_winner == 1, ]$genre_count)
```

```{r}
cor(d$best_picture_winner, d$genre_count)
```

```{r}
d %>%
  group_by(genre_count) %>%
  summarise(
    win_rate = mean(best_picture_winner == 1, na.rm = TRUE),
    n = n(),
    # Calculate Standard Error for a proportion
    se = sqrt((win_rate * (1 - win_rate)) / n)
  ) %>%
  mutate(
    # 95% CI calculation (1.96 is the z-score for 95%)
    lower = pmax(0, win_rate - 1.96 * se), # pmax(0,...) ensures we don't go below 0%
    upper = win_rate + 1.96 * se
  ) %>%
  ggplot(aes(x = factor(genre_count), y = win_rate)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  # Add the error bars
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, color = "darkred") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Number of Genres", 
    y = "Win Rate", 
    title = "Best Picture Win Rate by Genre Count",
    subtitle = "Red bars indicate 95% Confidence Intervals"
  )
```

# Correlation Map

```{r}
# 1. Filter for numeric variables AND exclude all genres
# This keeps the 'clean' numerical/award variables
numeric_data_no_genres <- d %>% 
  select(where(is.numeric)) %>% 
  select(-starts_with("genre_"))

# 2. Compute Correlation
corr_matrix <- cor(numeric_data_no_genres, use = "complete.obs")

# 3. Plot Heatmap
# Since we have fewer variables now, we can set lab = TRUE to see the exact coefficients
ggcorrplot(corr_matrix, 
           hc.order = TRUE, 
           type = "lower",
           lab = TRUE, 
           lab_size = 3,
           title = "Oscar Variable Correlation (Excluding Genres)",
           colors = c("#6D9EC1", "white", "#E46726"))

```

# Golden Globe Award Conversion Rate to Oscar Best Picture's Award 

## Testing for Golden Globe Award - Drama category
```{r}
# 1. Prep the data with the three categories and CI calculations
plot_data_gg <- d %>%
  ungroup() %>% 
  # Filter out rows where we don't know the nomination status
  filter(!is.na(nominee_drama_gg)) %>%
  mutate(gg_drama_status = case_when(
    nominee_drama_gg == 1 & won_drama_gg == 1 ~ "Won GG-Drama",
    nominee_drama_gg == 1 & won_drama_gg == 0 ~ "Nominated, No Win",
    TRUE ~ "Not Nominated"
  )) %>%
  # Set the order so the bars flow logically
  mutate(gg_drama_status = factor(gg_drama_status, 
                                  levels = c("Not Nominated", 
                                             "Nominated, No Win", 
                                             "Won GG-Drama"))) %>%
  # Note: Fixing the level name above to match the case_when
  mutate(gg_drama_status = factor(gg_drama_status, 
                                  levels = c("Not Nominated", 
                                             "Nominated, No Win", 
                                             "Won GG-Drama"))) %>%
  group_by(gg_drama_status) %>%
  summarise(
    win_rate = mean(as.numeric(best_picture_winner), na.rm = TRUE),
    n = n(),
    se = sqrt((win_rate * (1 - win_rate)) / n),
    .groups = "drop"
  ) %>%
  mutate(
    lower = pmax(0, win_rate - 1.96 * se),
    upper = pmin(1, win_rate + 1.96 * se)
  )

# 2. Plotting the results
ggplot(plot_data_gg, aes(x = gg_drama_status, y = win_rate, fill = gg_drama_status)) +
  geom_col(width = 0.7, alpha = 0.8) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, color = "darkred") +
  # Adding percentage labels above the bars
  geom_text(aes(y = upper, label = percent(win_rate, accuracy = 0.1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  scale_fill_manual(values = c("gray70", "lightcoral", "turquoise4")) +
  labs(
    title = "The Golden Globe Drama Ladder",
    subtitle = "Best Picture Win Rate by Golden Globe Recognition Level",
    x = "Golden Globe Drama Status",
    y = "Oscar Win Probability"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```
## Testing for Golden Globe Award - Musical/Comedy category

```{r}
# 1. Prep the data for Comedy (Corrected variable names)
plot_data_gg_comedy <- d %>%
  ungroup() %>% 
  filter(!is.na(nominee_comedy_gg)) %>%
  # Make sure we use ONE consistent name for the status column
  mutate(gg_comedy_status = case_when(
    nominee_comedy_gg == 1 & won_comedy_gg == 1 ~ "Won GG-Comedy",
    nominee_comedy_gg == 1 & won_comedy_gg == 0 ~ "Nominated, No Win",
    TRUE ~ "Not Nominated"
  )) %>%
  # Convert to factor using the CORRECT name we just created
  mutate(gg_comedy_status = factor(gg_comedy_status, 
                                  levels = c("Not Nominated", 
                                             "Nominated, No Win", 
                                             "Won GG-Comedy"))) %>%
  group_by(gg_comedy_status) %>%
  summarise(
    win_rate = mean(as.numeric(best_picture_winner), na.rm = TRUE),
    n = n(),
    se = sqrt((win_rate * (1 - win_rate)) / n),
    .groups = "drop"
  ) %>%
  mutate(
    lower = pmax(0, win_rate - 1.96 * se),
    upper = pmin(1, win_rate + 1.96 * se)
  )

# 2. Plotting the results
ggplot(plot_data_gg_comedy, aes(x = gg_comedy_status, y = win_rate, fill = gg_comedy_status)) +
  geom_col(width = 0.7, alpha = 0.8) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, color = "darkred") +
  geom_text(aes(y = upper, label = percent(win_rate, accuracy = 0.1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  scale_fill_manual(values = c("gray70", "lightgoldenrod3", "springgreen4")) + # Different colors for Comedy
  labs(
    title = "The Golden Globe Musical/Comedy Ladder",
    subtitle = "Best Picture Win Rate by Golden Globe Recognition Level",
    x = "Golden Globe Musical/Comedy Status",
    y = "Oscar Win Probability"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

## Understanding the actual number of Golden Globe winner that won Oscars' Best Picture
```{r}
# Create a summary for Drama
drama_summary <- d %>%
  group_by(won_drama_gg) %>%
  summarise(
    Total_Films = n(),
    Oscar_Wins = sum(best_picture_winner),
    Win_Rate = mean(best_picture_winner)
  ) %>%
  mutate(Category = "Drama") %>%
  rename(Won_GG = won_drama_gg)

# Create a summary for Comedy
comedy_summary <- d %>%
  group_by(won_comedy_gg) %>%
  summarise(
    Total_Films = n(),
    Oscar_Wins = sum(best_picture_winner),
    Win_Rate = mean(best_picture_winner)
  ) %>%
  mutate(Category = "Comedy") %>%
  rename(Won_GG = won_comedy_gg)

# Combine them into one professional table
gg_conversion_table <- bind_rows(drama_summary, comedy_summary) %>%
  filter(Won_GG == 1) %>% # We only care about the films that actually won the GG
  select(Category, Total_Films, Oscar_Wins, Win_Rate) %>%
  mutate(Win_Rate = scales::percent(Win_Rate, accuracy = 0.1))

print(gg_conversion_table)

```

# Does winning for best editing film convert to winning Oscar Best Picture

```{r}
# 1. Prep the data with CI calculations
plot_data <- d %>%
  ungroup() %>% 
  mutate(editing_status = case_when(
    nominee_film_edit == 1 & award_film_edit == 1 ~ "Won Best Editing",
    nominee_film_edit == 1 & award_film_edit == 0 ~ "Nominated, No Win",
    TRUE ~ "Not Nominated"
  )) %>%
  mutate(editing_status = factor(editing_status, 
                                 levels = c("Not Nominated", 
                                            "Nominated, No Win", 
                                            "Won Best Editing"))) %>%
  group_by(editing_status) %>%
  summarise(
    win_rate = mean(as.numeric(best_picture_winner), na.rm = TRUE),
    n = n(),
    # Calculate Standard Error
    se = sqrt((win_rate * (1 - win_rate)) / n),
    .groups = "drop"
  ) %>%
  mutate(
    # Calculate bounds
    lower = pmax(0, win_rate - 1.96 * se),
    upper = pmin(1, win_rate + 1.96 * se)
  )

# 2. Plot with Error Bars
ggplot(plot_data, aes(x = editing_status, y = win_rate, fill = editing_status)) +
  geom_bar(stat = "identity", show.legend = FALSE, alpha = 0.9) +
  # Add the error bars
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, color = "darkred") +
  # Adjust text label position to be above the error bar for clarity
  geom_text(aes(y = upper, label = percent(win_rate, accuracy = 0.1)), 
            vjust = -0.8, size = 4) + 
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  scale_fill_manual(values = c("gray70", "steelblue3", "dodgerblue4")) +
  labs(title = "The 'Best Editing' Prerequisite",
       subtitle = "95% Confidence Intervals for Best Picture Win Rate",
       x = "Film Editing Recognition Level",
       y = "Best Picture Win Rate") +
  theme_light()
```

# The distribution of winners in box office and tomotometer

```{r}
# Tomatometer vs Box Office
ggplot(d, aes(x = tomatometer, y = box_office_norm, color = factor(best_picture_winner))) +
  geom_point(aes(size = total_academy_nominations), alpha = 0.6) +
  scale_color_manual(values = c("grey", "gold"), name = "Best Picture Winner") +
  labs(title = "Critic Scores vs. Commercial Success",
       x = "Tomatometer Score",
       y = "Box Office (Normalized-Year)") +
  theme_bw()
```


# The distribution of winners in box office and tomotometer

```{r}
# Tomatometer vs Popcornmeter
ggplot(d, aes(x = tomatometer, y = popcornmeter, color = factor(best_picture_winner))) +
  geom_point(aes(size = total_academy_nominations), alpha = 0.6) +
  scale_color_manual(values = c("grey", "gold"), name = "Best Picture Winner") +
  labs(title = "Critic Scores Public VS Expert",
       x = "Tomatometer Score",
       y = "Popcornmeter Score") +
  theme_bw()

```

```{r}
d[d$popcornmeter<25,]
```

```{r}
# 1. Calculate average Tomatometer and CI
tomatometer_stats <- d %>%
  filter(!is.na(tomatometer)) %>%
  group_by(best_picture_winner) %>%
  summarise(
    avg_score = mean(tomatometer, na.rm = TRUE),
    n = n(),
    sd_val = sd(tomatometer, na.rm = TRUE),
    se = sd_val / sqrt(n),
    .groups = "drop"
  ) %>%
  mutate(
    lower = avg_score - 1.96 * se,
    upper = avg_score + 1.96 * se
  )

# 2. Plotting
ggplot(tomatometer_stats, aes(x = factor(best_picture_winner), y = avg_score, fill = factor(best_picture_winner))) +
  geom_col(width = 0.6, alpha = 0.8) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, color = "darkred") +
  scale_x_discrete(labels = c("0" = "Non-Winner", "1" = "Winner")) +
  scale_fill_manual(values = c("gray70", "#FF6347")) + # Tomato-themed color
  labs(
    title = "Average Tomatometer Score by Winner Status",
    subtitle = "Error bars indicate 95% Confidence Intervals",
    x = "Award Outcome",
    y = "Average Tomatometer Score (%)",
    fill = "Result"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
# 1. Calculate average Popcornmeter and CI
popcorn_stats <- d %>%
  filter(!is.na(popcornmeter)) %>%
  group_by(best_picture_winner) %>%
  summarise(
    avg_score = mean(popcornmeter, na.rm = TRUE),
    n = n(),
    sd_val = sd(popcornmeter, na.rm = TRUE),
    se = sd_val / sqrt(n),
    .groups = "drop"
  ) %>%
  mutate(
    lower = avg_score - 1.96 * se,
    upper = avg_score + 1.96 * se
  )

# 2. Plotting
ggplot(popcorn_stats, aes(x = factor(best_picture_winner), y = avg_score, fill = factor(best_picture_winner))) +
  geom_col(width = 0.6, alpha = 0.8) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, color = "darkred") +
  scale_x_discrete(labels = c("0" = "Non-Winner", "1" = "Winner")) +
  scale_fill_manual(values = c("gray70", "#FFD700")) + # Golden/Popcorn color
  labs(
    title = "Average Popcornmeter Score by Winner Status",
    subtitle = "Error bars indicate 95% Confidence Intervals",
    x = "Award Outcome",
    y = "Average Popcornmeter Score (%)",
    fill = "Result"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```



# Inspecting distribution of variables across winners and non-winners group
## These distributions are very similar between winning and non-winning group
```{r}
# Runtime distribution
ggplot(d, aes(x = running_time_minutes, fill = factor(best_picture_winner))) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Film Length", x = "Minutes", fill = "Winner") +
  theme_minimal()
```

```{r}
# Words count
ggplot(d, aes(x = words, fill = factor(best_picture_winner))) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Words Count", x = "Word Count", fill = "Winner") +
  theme_minimal()
```

```{r}
# Character count
ggplot(d, aes(x = characters, fill = factor(best_picture_winner))) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Character Number", x = "Character Count", fill = "Winner") +
  theme_minimal()
```

# Inspecting the distribution of award nominations and the winning rate 

```{r}

ggplot(d, aes(x = total_academy_nominations, fill = factor(best_picture_winner))) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Character Number", x = "Character Count", fill = "Winner") +
  theme_minimal()
```

```{r}
# 1. Calculate average nominations and CI
nomination_stats <- d %>%
  group_by(best_picture_winner) %>%
  summarise(
    avg_noms = mean(total_academy_nominations, na.rm = TRUE),
    n = n(),
    sd_val = sd(total_academy_nominations, na.rm = TRUE),
    se = sd_val / sqrt(n),
    .groups = "drop"
  ) %>%
  mutate(
    lower = pmax(0, avg_noms - 1.96 * se),
    upper = avg_noms + 1.96 * se
  )

# 2. Plotting
ggplot(nomination_stats, aes(x = factor(best_picture_winner), y = avg_noms, fill = factor(best_picture_winner))) +
  geom_col(width = 0.6, alpha = 0.8) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, color = "darkred") +
  scale_x_discrete(labels = c("0" = "Non-Winner", "1" = "Winner")) +
  scale_fill_manual(values = c("gray70", "#E7B800")) +
  labs(
    title = "Average Academy Nominations by Winner Status",
    subtitle = "Error bars indicate 95% Confidence Intervals",
    x = "Award Outcome",
    y = "Average Total Nominations",
    fill = "Result"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

# Inspecting the distribution of award won and the winning rate 

```{r}
# Number of Awards excluding Best Pictures
ggplot(d, aes(x = total_academy_award_exclude_best_pic, fill = factor(best_picture_winner))) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Awards Exclude Best Pics", x = "Total Academy Wins Exclude Best Pics", fill = "Winner") +
  theme_minimal()
```

```{r}
# 1. Calculate average wins and CI
award_win_stats <- d %>%
  group_by(best_picture_winner) %>%
  summarise(
    avg_wins = mean(total_academy_award_exclude_best_pic, na.rm = TRUE),
    n = n(),
    sd_val = sd(total_academy_award_exclude_best_pic, na.rm = TRUE),
    se = sd_val / sqrt(n),
    .groups = "drop"
  ) %>%
  mutate(
    lower = pmax(0, avg_wins - 1.96 * se),
    upper = avg_wins + 1.96 * se
  )

# 2. Plotting
ggplot(award_win_stats, aes(x = factor(best_picture_winner), y = avg_wins, fill = factor(best_picture_winner))) +
  geom_col(width = 0.6, alpha = 0.8) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, color = "darkred") +
  scale_x_discrete(labels = c("0" = "Non-Winner", "1" = "Winner")) +
  scale_fill_manual(values = c("gray70", "#00AFBB")) +
  labs(
    title = "Average Academy Wins (Excl. Best Picture) by Winner Status",
    subtitle = "Error bars indicate 95% Confidence Intervals",
    x = "Award Outcome",
    y = "Average Wins (Excluding Best Picture)",
    fill = "Result"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

```{r}
# Normalized Box Office
ggplot(d, aes(x = box_office_norm, fill = factor(best_picture_winner))) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Normalized Box Office (By Year)", x = "Total Academy Wins Exclude Best Pics", fill = "Winner") +
  theme_minimal()
```

```{r}
# Normalized Budget
d <- d %>%
  group_by(release_year) %>%
  mutate(budget_norm = budget_millions / sum(budget_millions, na.rm = TRUE)) %>%
  ungroup()


ggplot(d, aes(x = budget_norm, fill = factor(best_picture_winner))) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Normalized Budget (By Year)", x = "Total Academy Wins Exclude Best Pics", fill = "Winner") +
  theme_minimal()
```

```{r}
# Normalized Surplus
# Isolate inflation/price shocks

d$surplus_norm <- d$box_office_norm - d$budget_norm

ggplot(d, aes(x = surplus_norm, fill = factor(best_picture_winner))) +
  geom_density(alpha = 0.4) +
  labs(title = "Distribution of Normalized Surplus (By Year)", x = "Total Academy Wins Exclude Best Pics", fill = "Winner") +
  theme_minimal()
```


# Now, I want to verify
1. Can we use PCA to move them into different clusters
2. If we can separate them into different clusters, what factors determine them the most

```{r}
# 1. Prepare the data, left out the genre variables
pca_data <- d %>% 
  select(where(is.numeric)) %>% 
  select(-starts_with("genre_")) %>% 
  select(-"box_office_norm")%>% 
  select(-"best_picture_winner") #Excluding response variable

# 2. Perform PCA
# 'scale = TRUE' is critical because 'budget' is in millions while 'tomatometer' is 0-100
res.pca <- prcomp(pca_data, scale = TRUE)

# 3. Visualization: Individual Films & Variables (Biplot)
fviz_pca_biplot(res.pca, 
                label = "var",        # Only label variables to avoid cluttering with film names
                habillage = as.factor(d$best_picture_winner), # Color by Win Status
                addEllipses = TRUE,   # Concentration ellipses to show cluster separation
                col.var = "black",    # Color of variable arrows
                alpha.ind = 0.5) +    # Transparency for individual films
  labs(title = "PCA: Oscar Winner Clusters",
       subtitle = "Variables contributing to Component 1 and 2") +
  theme_minimal()

# 4. Visualization: Contribution of Variables
fviz_pca_var(res.pca, col.var = "contrib", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE) # Avoid overlapping labels

```

# Examine the relationship between directors history/prestige and the winning status of the film

```{r}
# 1. Define the variables
dir_vars <- c("best_director_award_before", 
              "best_director_nominee_before", 
              "best_director_award_now", 
              "best_director_nominee_include_now")

# 2. Calculate the stats with Confidence Intervals
impact_df <- map_df(dir_vars, function(v) {
  d %>%
    filter(!!sym(v) == 1) %>% 
    summarise(
      variable = v,
      win_rate = mean(best_picture_winner, na.rm = TRUE),
      n = n(),
      # Calculate Standard Error for the proportion
      se = sqrt((win_rate * (1 - win_rate)) / n)
    )
}) %>%
  mutate(
    # 95% CI Calculation
    lower = pmax(0, win_rate - 1.96 * se),
    upper = pmin(1, win_rate + 1.96 * se),
    # Labels
    clean_label = case_when(
      variable == "best_director_award_now" ~ "Won Best Director (This Year)",
      variable == "best_director_nominee_include_now" ~ "Nominated for Director (This Year)",
      variable == "best_director_award_before" ~ "Has Won BD Previously",
      variable == "best_director_nominee_before" ~ "Has Been Nominated Previously",
      TRUE ~ variable
    ),
    context = if_else(str_detect(variable, "now"), "Current Momentum", "Past Pedigree")
  )

# 3. Create the Visualization with Error Bars
ggplot(impact_df, aes(x = reorder(clean_label, win_rate), y = win_rate, fill = context)) +
  geom_col(alpha = 0.7, width = 0.7) +
  # Adding the error bars
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, color = "darkred", alpha = 0.8) +
  # Move labels slightly further out to avoid the error bars
  geom_text(aes(y = upper, label = percent(win_rate, accuracy = 0.1)), 
            hjust = -0.3, fontface = "bold", size = 3.5) +
  coord_flip() +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  scale_fill_manual(values = c("Current Momentum" = "#E7B800", "Past Pedigree" = "#00AFBB")) +
  labs(
    title = "Directorial Impact on Best Picture Success",
    subtitle = "Error bars indicate 95% Confidence Intervals",
    x = NULL,
    y = "Probability of Winning Best Picture",
    fill = "Context"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    axis.text.y = element_text(face = "bold", size = 10)
  )
```

# Investigating rating VS winning status

```{r}
table(d$rating, d$best_picture_winner)
plot(table(d$best_picture_winner,d$rating))
```

```{r}
# As we increase from G to R, it means there is more restriction to the film
cor(d$best_picture_winner, as.numeric(factor(d$rating)))
```

```{r}
# Plot the bar graph
table(d$rating)
```






