---
title: "Life Expectancy"
author: "Henry Tan, Casper Liao"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, #displaying actual code
                      error = TRUE, # still knit even with errors
                      tidy.opts=list(width.cutoff=80), 
                      tidy  = TRUE, # tidy up spacing for R code
                      eval = TRUE, collapse = TRUE, comment = NA)

# collapse = TRUE, code and result tucked together into one cohesive grey box
# comment = NA, normally add ## before the output
```

# Problem Statment
To what extent that the socioeconomic factors could affect individual life expectancy in the US?

# Dataset Description and Document
The dataset is obtained from this github repo: https://github.com/mi3nts/life-expectancy.
The author collected and compiled life expectancy and its corresponding demographic data according to different counties, from 2012 to 2019. 

The data is collected and combined from the three sources below:

(1) Demographic and socioeconomic data from US Census and related sources

(2) Atmospheric reanalysis data from CAMS (Copernicus Atmosphere Monitoring Service)

(3) Environmental variables from ECMWF ERA5 reanalysis

The original dataset contains 15 socioeconomic features and 87 environmental variables. However, since our question only focus on the impact of socioeconomic factors, we will just use the 15 socioeconomic features for our analysis. 

The dataset contains data from 2012 to 2019. For our preliminary analysis, we will only focus on **2019** data. 

Full description about the data columns are available in the same github repo: https://github.com/mi3nts/life-expectancy. 

# Loading Dataset

```{r}
# Full dataset with socioeconomic + environmental features
full.data <- read.csv("final_combined_all_variables.csv")
dim(full.data)
```


```{r}
# Filter the data
# Extract only the first 19 socioeconomic variables and 2019 data
row.extract <- which(full.data$Year == 2019)
life.data <- full.data[row.extract,1:19] 
```

```{r}
# Clean up the long column name
colnames(life.data) <- c(
  "county", "state", "fips",                 # Identifiers
  "life_exp",                                # Target
  "poverty_rate", "edu_hs", "edu_bach",      # Socioeconomic
  "median_age", "gini_index", "disability", 
  "total_pop", "median_income", "unemployment", 
  "pop_white", "pop_hispanic", "pop_black",
  "no_vehicle", "rent_burden", "single_mother" # Housing/Transport
)

# Quick check to ensure no weird symbols remain
colnames(life.data) <- make.names(colnames(life.data))
```

```{r}
# Preview your clean names
head(life.data)
dim(life.data)
```


```{r}
# Inspect the structure of the data
str(life.data)
```


```{r}
# Export the dataset
write.csv(life.data, "2019_life_socio.csv",row.names = FALSE)
```










