---
title: "model"
output: html_document
---

# We create models

# combine golden globe awards (merging drama and comedy)
```{r}
dat <- read.csv("academy_final.csv")
dat$total_nominee_gg <- dat$nominee_drama_gg + dat$nominee_comedy_gg
dat$total_won_gg <- dat$won_drama_gg + dat$won_comedy_gg

dat$nominee_drama_gg <- NULL
dat$nominee_comedy_gg <- NULL
dat$won_drama_gg <- NULL
dat$won_comedy_gg <- NULL
```

# categorize and combine genre: too many of them, we eliminate them
```{r}
# New genre group:
# 1. Drama
# 2. Human Interest (Biography, History, Romance, LGBTQ, War, Music)
# 3. Suspense (Mystery & Thriller, Crime, Sci-Fi, Fantasy, Adventure)
# 4. Entertainment (Action, Western, Comedy, Musical)
# 5. Niche (Animation, Holiday, Horror, Kids & Family, Sports)

# Define which original columns map to each new group
human_interest <- c("genre_Biography", "genre_History", "genre_Romance", 
                    "genre_LGBTQ_", "genre_War", "genre_Music")

suspense       <- c("genre_Mystery___Thriller", "genre_Crime", "genre_Sci_Fi", 
                    "genre_Fantasy", "genre_Adventure")

entertainment  <- c("genre_Action", "genre_Western", "genre_Comedy", "genre_Musical")

niche          <- c("genre_Animation", "genre_Holiday", "genre_Horror", 
                    "genre_Kids___Family", "genre_Sports")

# Create new binary indicators (1 if ANY of the group's genres apply)
dat$genre_Drama         <- dat$genre_Drama  # already exists, keep as-is
dat$genre_HumanInterest <- as.integer(rowSums(dat[, human_interest]) > 0)
dat$genre_Suspense      <- as.integer(rowSums(dat[, suspense]) > 0)
dat$genre_Entertainment <- as.integer(rowSums(dat[, entertainment]) > 0)
dat$genre_Niche         <- as.integer(rowSums(dat[, niche]) > 0)

# Drop all original genre_* columns
original_genres <- c("genre_Comedy", "genre_History", "genre_Biography",
                     "genre_Mystery___Thriller", "genre_Romance", "genre_Adventure",
                     "genre_Action", "genre_Crime", "genre_Fantasy", "genre_Sci_Fi",
                     "genre_War", "genre_LGBTQ_", "genre_Music", "genre_Musical",
                     "genre_Western", "genre_Horror", "genre_Kids___Family",
                     "genre_Animation", "genre_Holiday", "genre_Sports")

dat <- dat[, !names(dat) %in% original_genres]
str(dat)
```

# go through the data set, go through and inspect binary predictor that are 95% zeros
```{r}
binary_cols <- names(dat)[sapply(dat, function(x) all(x %in% c(0, 1, NA)))]

zero_pct <- sapply(dat[, binary_cols], function(x) mean(x == 0, na.rm = TRUE))

flagged <- sort(zero_pct[zero_pct >= 0.95], decreasing = TRUE)
flagged

# only genre_Niche, bye bye!
dat$genre_Niche <- NULL
```

# drop 
1. film
2. director
3. release year
4. genre
5. wikipedia
6. rotten_tomato
7. language
```{r}
dat$post_2009 <- as.integer(dat$release_year >= 2009)

drop_cols <- c("film", "director", "release_year", "genre", 
               "wikipedia", "rotten_tomato", "language")
dat <- dat[, !names(dat) %in% drop_cols]
dat$post_2009
```
# Tidying up rating variables
```{r}
# extract base rating (e.g. "PG-13", "R", "PG") from full rating string
dat$rating <- gsub(" \\(.*", "", dat$rating)
table(dat$rating)

# filling NA value
dat[which(is.na(dat$rating)),]

# Film "Drive My Car" is "not rated", we classify it as R
dat[which(is.na(dat$rating)),]$rating <- "R"
table(dat$rating)
```

# start with our best guess model with linear regression
```{r}
model.sig <- lm(best_picture_winner ~ total_academy_award_exclude_best_pic + 
                                   total_academy_nominations + 
                                   nominee_film_edit + 
                                   award_film_edit + 
                                   popcornmeter, 
            data = dat)

summary(model.sig)
par(mfrow=c(2,2))
plot(model.sig)
```

# linear model appears to be not suitable --> try logistic regression model
```{r}
model.sig.logit <- glm(best_picture_winner ~ total_academy_award_exclude_best_pic + 
                                          total_academy_nominations + 
                                          nominee_film_edit + 
                                          award_film_edit + 
                                          popcornmeter, 
                   data = dat, 
                   family = binomial)

summary(model.sig.logit)
```

```{r}
library(arm)
library(car)
library(ggplot2)
library(pROC)

# 1. Binned Residual Plot - want ~95% of points within grey bands
binnedplot(fitted(model.sig.logit), residuals(model.sig.logit, type = "response"),
           main = "Binned Residual Plot")

# 2. VIF - check multicollinearity, want < 5
vif(model.sig.logit)

# 3. Linearity of log-odds for continuous predictors
dat$log_odds <- predict(model.sig.logit, type = "link")
ggplot(dat, aes(x = popcornmeter, y = log_odds)) + geom_point() + geom_smooth()
ggplot(dat, aes(x = total_academy_award_exclude_best_pic, y = log_odds)) + geom_point() + geom_smooth()
ggplot(dat, aes(x = total_academy_nominations, y = log_odds)) + geom_point() + geom_smooth()

# 4. Influential points via Cook's distance
plot(model.sig.logit, which = 4)

# 5. ROC/AUC - model discrimination, want > 0.8
roc_curve <- roc(dat$best_picture_winner, fitted(model.sig.logit))
plot(roc_curve, main = "ROC Curve")
auc(roc_curve)

```
# check for individual predictor contribution
```{r}
# Individual logistic regression for each non-significant predictor

summary(glm(best_picture_winner ~ total_academy_nominations, data = dat, family = binomial))

summary(glm(best_picture_winner ~ award_film_edit, data = dat, family = binomial))

summary(glm(best_picture_winner ~ popcornmeter, data = dat, family = binomial))
```

```{r}
# academy nomination + awards
summary(glm(best_picture_winner ~ total_academy_nominations + total_academy_award_exclude_best_pic, data = dat, family = binomial))

# film editing nomination + awards
summary(glm(best_picture_winner ~ award_film_edit + nominee_film_edit, data = dat, family = binomial))

# film editing nomination + awards + total_academy_award_exclude_best_pic
summary(glm(best_picture_winner ~ award_film_edit + nominee_film_edit + total_academy_award_exclude_best_pic, data = dat, family = binomial))
```

# new base model: 
# 1. total_academy_award_exclude_best_pic
# 2. nominee_film_edit
```{r}
summary(glm(best_picture_winner ~ nominee_film_edit + total_academy_award_exclude_best_pic, data = dat, family = binomial))
```
# add in secondary significance variables
```{r}
# secondary model: base + genre + golden globe variables
summary(glm(best_picture_winner ~ nominee_film_edit + 
                                   total_academy_award_exclude_best_pic +
                                   genre_Drama +
                                   genre_HumanInterest +
                                   genre_Suspense +
                                   genre_Entertainment +
                                   total_nominee_gg +
                                   total_won_gg, 
            data = dat, family = binomial))
```

# interactions with post_2009
```{r}
summary(glm(best_picture_winner ~ nominee_film_edit + 
                                   total_academy_award_exclude_best_pic +
                                   genre_Drama +
                                   genre_HumanInterest +
                                   genre_Suspense +
                                   genre_Entertainment +
                                   genre_Drama * post_2009 +
                                   genre_HumanInterest * post_2009 +
                                   genre_Suspense * post_2009 +
                                   genre_Entertainment * post_2009, 
            data = dat, family = binomial))

summary(glm(best_picture_winner ~ total_academy_award_exclude_best_pic + total_academy_award_exclude_best_pic * post_2009, data = dat, family = binomial))


summary(glm(best_picture_winner ~ nominee_film_edit + nominee_film_edit * post_2009, data = dat, family = binomial))
```


```{r}

table(dat$nominee_film_edit, dat$post_2009, dat$best_picture_winner)
```










