---
title: "Statistical Model Fitting"
output: pdf_document
---

# This file contains codes that produce all model-fitting related visuals in this project. 


# Feature Engineering

## 1. Combine golden globe awards (merging drama and comedy)
```{r}
dat <- read.csv("academy_final.csv")
dat$total_nominee_gg <- dat$nominee_drama_gg + dat$nominee_comedy_gg
dat$total_won_gg <- dat$won_drama_gg + dat$won_comedy_gg

dat$nominee_drama_gg <- NULL
dat$nominee_comedy_gg <- NULL
dat$won_drama_gg <- NULL
dat$won_comedy_gg <- NULL
```

## 2. Categorize and combine genre: too many of them, we eliminate them
```{r}
# New genre group:
# 1. Drama
# 2. Human Interest (Biography, History, Romance, LGBTQ, War, Music)
# 3. Suspense (Mystery & Thriller, Crime, Sci-Fi, Fantasy, Adventure)
# 4. Entertainment (Action, Western, Comedy, Musical)
# 5. Niche (Animation, Holiday, Horror, Kids & Family, Sports)

# Define which original columns map to each new group
human_interest <- c("genre_Biography", "genre_History", "genre_Romance", 
                    "genre_LGBTQ_", "genre_War", "genre_Music")

suspense       <- c("genre_Mystery___Thriller", "genre_Crime", "genre_Sci_Fi", 
                    "genre_Fantasy", "genre_Adventure")

entertainment  <- c("genre_Action", "genre_Western", "genre_Comedy", "genre_Musical")

niche          <- c("genre_Animation", "genre_Holiday", "genre_Horror", 
                    "genre_Kids___Family", "genre_Sports")

# Create new binary indicators (1 if ANY of the group's genres apply)
dat$genre_Drama         <- dat$genre_Drama  # already exists, keep as-is
dat$genre_HumanInterest <- as.integer(rowSums(dat[, human_interest]) > 0)
dat$genre_Suspense      <- as.integer(rowSums(dat[, suspense]) > 0)
dat$genre_Entertainment <- as.integer(rowSums(dat[, entertainment]) > 0)
dat$genre_Niche         <- as.integer(rowSums(dat[, niche]) > 0)

# Drop all original genre_* columns
original_genres <- c("genre_Comedy", "genre_History", "genre_Biography",
                     "genre_Mystery___Thriller", "genre_Romance", "genre_Adventure",
                     "genre_Action", "genre_Crime", "genre_Fantasy", "genre_Sci_Fi",
                     "genre_War", "genre_LGBTQ_", "genre_Music", "genre_Musical",
                     "genre_Western", "genre_Horror", "genre_Kids___Family",
                     "genre_Animation", "genre_Holiday", "genre_Sports")

dat <- dat[, !names(dat) %in% original_genres]
str(dat)
```
```{r}
sum(dat$genre_Drama)
sum(dat$genre_HumanInterest)
sum(dat$genre_Suspense)
sum(dat$genre_Entertainment)
sum(dat$genre_Niche)
```
## 3. Normalize budget and box office
```{r}
library(dplyr)
# normalize box office and budget by year (run before dropping release_year)
dat <- dat %>%
  group_by(release_year) %>%
  mutate(
    box_office_norm = box_office_millions / sum(box_office_millions, na.rm = TRUE),
    budget_norm = budget_millions / sum(budget_millions, na.rm = TRUE)
  ) %>%
  ungroup()

dat$box_office_millions <- NULL
dat$budget_millions <- NULL
```

## 4. Tidying up rating variables
```{r}
# extract base rating (e.g. "PG-13", "R", "PG") from full rating string
dat$rating <- gsub(" \\(.*", "", dat$rating)
table(dat$rating)
# filling NA value
dat[which(is.na(dat$rating)),]
# Film "Drive My Car" is "not rated", we classify it as R
dat[which(is.na(dat$rating)),]$rating <- "R"
table(dat$rating)

# collapse G and PG into one category to avoid separation issues (only 1 G film)
dat$rating <- ifelse(dat$rating %in% c("G", "PG"), "PG_under", dat$rating)
dat$rating <- factor(dat$rating, levels = c("PG_under", "PG-13", "R"))
table(dat$rating)
```

## 5. Go through the data set, go through and inspect binary predictor that are 95% zeros
```{r}
binary_cols <- names(dat)[sapply(dat, function(x) all(x %in% c(0, 1, NA)))]

zero_pct <- sapply(dat[, binary_cols], function(x) mean(x == 0, na.rm = TRUE))

flagged <- sort(zero_pct[zero_pct >= 0.95], decreasing = TRUE)
flagged

# only genre_Niche, bye bye!
dat$genre_Niche <- NULL
```

## 6. Dropping unnecessary/extra variables before model fitting 
1. film
2. director
3. release year
4. genre
5. wikipedia
6. rotten_tomato
7. language

## 7. Creating extra variable called 2009 to examine the effect of major Oscar change

```{r}
dat$post_2009 <- as.integer(dat$release_year >= 2009)

drop_cols <- c("film", "director", "release_year", "genre", 
               "wikipedia", "rotten_tomato", "language")
dat <- dat[, !names(dat) %in% drop_cols]
dat$post_2009
```

## 8. Inspecting the dataset now
```{r}
dat <- data.frame(dat)
str(dat)
```

# Model Fitting

## 1. Start with our best guess model with linear regression
```{r}
model.sig <- lm(best_picture_winner ~ total_academy_award_exclude_best_pic + 
                                   total_academy_nominations + 
                                   nominee_film_edit + 
                                   award_film_edit + 
                                   popcornmeter, data = dat)

par(mfrow=c(2,2))
plot(model.sig)
```

## 2. Clear violation of linear regression assumptions, switch to logistic regression

```{r}
model.sig.logit <- glm(best_picture_winner ~ total_academy_award_exclude_best_pic + 
                                          total_academy_nominations + 
                                          nominee_film_edit + 
                                          award_film_edit + 
                                          popcornmeter, 
                   data = dat, 
                   family = binomial)

summary(model.sig.logit)
vif(model.sig.logit)
```

## This checks if the model is performing good in residuals/any multicollinear - pretty good
```{r}
# Install if you haven't yet
# install.packages("performance")
# install.packages("see") 

library(performance)
library(see) # Makes the plots look professional

check_model(model.sig.logit)
```

## This checks if the continuous variables are linear with respect to log_odds
If the continuous variables are linear with respect to log_odds, the residuals should be distributed evenly across different values, and the indicator line should appear flat. We fulfill the assumption check. 
```{r}
library(car)
residualPlots(model.sig.logit)
```

## Check for individual predictor contribution
We suspect the total_academy_nominations and award_film_edit are important too, and their information might be masked by the other variables included.

```{r}
# Individual logistic regression for each non-significant predictor

summary(glm(best_picture_winner ~ total_academy_nominations, data = dat, family = binomial))

summary(glm(best_picture_winner ~ award_film_edit, data = dat, family = binomial))

summary(glm(best_picture_winner ~ popcornmeter, data = dat, family = binomial))
```

```{r}
# academy nomination + awards
summary(glm(best_picture_winner ~ total_academy_nominations + total_academy_award_exclude_best_pic, data = dat, family = binomial))

# film editing nomination + awards
summary(glm(best_picture_winner ~ award_film_edit + nominee_film_edit, data = dat, family = binomial))

# film editing nomination + awards + total_academy_award_exclude_best_pic
summary(glm(best_picture_winner ~ award_film_edit + nominee_film_edit + total_academy_award_exclude_best_pic, data = dat, family = binomial))
```
At this stage, we know that the total academy awards won and nominated for best editing film did increase the probability of winning the Oscar's award (significant effect). Let's try adding another set of variables (the group of variables which we think might not be that significant). 

# Add in second group of significance variables
```{r}
# secondary model: base + genre + golden globe variables
second.model <- glm(best_picture_winner ~  nominee_film_edit +
                                   total_academy_nominations + 
                                   award_film_edit + 
                                   popcornmeter +
                                   total_academy_award_exclude_best_pic +
                                   genre_Drama +
                                   genre_HumanInterest +
                                   genre_Suspense +
                                   genre_Entertainment +
                                   total_nominee_gg +
                                   total_won_gg, data = dat, family = binomial)

summary(second.model)
vif(second.model)
```


```{r}
# We are trying to see if the addition of extra variables are important
# We use anova for likelihood ratio test (using chi square)
anova(model.sig.logit, second.model, test = "Chisq")
```
The result is insignificant, hence adding the additional variables above does not improve the model/add extra information. The coefficient of the extra variables added are statistically insignificant.

We can verify with AIC and BIC
```{r}
# Compare AIC for both models
AIC(model.sig.logit, second.model)

# Compare BIC for both models
BIC(model.sig.logit, second.model)
```
# Base group + add in the variables we think are insignificant

```{r}
third.model <- glm(best_picture_winner ~ total_academy_award_exclude_best_pic + 
                                          total_academy_nominations +
                                          nominee_film_edit + 
                                          award_film_edit + 
                                          popcornmeter + 
                                          box_office_norm +
                                          tomatometer +
                                          rating +
                                          best_director_award_now + 
                                          best_director_award_before +
                                          best_director_nominee_include_now +
                                          best_director_nominee_before +  # director history
                                          budget_norm +
                                          running_time_minutes +
                                          characters +                     # title name length
                                          words, data = dat, family = binomial)


vif(third.model)
```
There is extreme collinearity between some variables. We suspect it is:
1. best_director_nominee_include_now & best_director_nominee_before
2. characters and words

In addition, we realize that in rating, none of the categories under PG 13 win the Oscars. This might cause the perfect separation problem to the model, where film with PG 13 will be immediately predicted to lose the Oscars. We tried excluding rating and the standard error of intercept return to normal and smaller number. 

Let's remove and reconstruct the model
```{r}
third.model <- glm(best_picture_winner ~ total_academy_award_exclude_best_pic + 
                                          total_academy_nominations +
                                          nominee_film_edit + 
                                          award_film_edit + 
                                          popcornmeter + 
                                          box_office_norm +
                                          tomatometer +
                                          best_director_award_now + 
                                          best_director_award_before +
                                          best_director_nominee_include_now + # director history
                                          budget_norm +
                                          running_time_minutes +
                                          characters +
                                          post_2009, data = dat, family = binomial)


vif(third.model)
```
All VIF is lower than 5, we are good.

```{r}
summary(third.model)
vif(third.model)
```


```{r}
# We are trying to see if the addition of extra variables are important
# We use anova for likelihood ratio test (using chi square)
anova(model.sig.logit, third.model, test = "Chisq")
```
The addition of these new variables does not seem to contribute to extra information, but the probability is very close to significant 0.05, so I suspect there might be something useful in the extra variables but is masked by other noise.

```{r}
# Backward stepwise selection with AIC criteria
# This will automatically find the best combination of variables
refined_model <- step(third.model, direction = "backward")
summary(refined_model)
```


```{r}
# Interested to see if we will have the same variables being selected using full model
full.model <- glm(formula = best_picture_winner~.-rating, data=dat, family=binomial)
refined_model <- step(full.model, direction = "backward")
summary(refined_model)
```


# Interaction Term Inspection

## First interaction term: total_academy_nominations * genre_Drama
Not significant
```{r}
# Now I am interested to test if the addition of interaction term is important
final.model <- glm(formula = best_picture_winner ~ popcornmeter + total_academy_award_exclude_best_pic + 
                   nominee_film_edit + best_director_award_before + best_director_award_now + 
                   total_won_gg + budget_norm, family = binomial, data = dat)

# Adding interaction term
interact.model.1 <- glm(formula = best_picture_winner ~ popcornmeter + total_academy_award_exclude_best_pic + 
                   nominee_film_edit + best_director_award_before + best_director_award_now + 
                   total_won_gg + budget_norm + total_academy_nominations*genre_Drama, family = binomial, data = dat)

# Anova
anova(final.model, interact.model.1, test = "Chisq")
```

## Second interaction term: box_office_norm * budget_norm
Not significant
```{r}
# Now I am interested to test if the addition of interaction term is important
final.model <- glm(formula = best_picture_winner ~ popcornmeter + total_academy_award_exclude_best_pic + 
                   nominee_film_edit + best_director_award_before + best_director_award_now + 
                   total_won_gg + budget_norm, family = binomial, data = dat)

vif(final.model)
```

```{r}
# Adding interaction term
interact.model.2 <- glm(formula = best_picture_winner ~ popcornmeter + total_academy_award_exclude_best_pic + 
                   nominee_film_edit + best_director_award_before + best_director_award_now + 
                   total_won_gg + budget_norm + box_office_norm * budget_norm, family = binomial, data = dat)

# Anova
anova(final.model, interact.model.2, test = "Chisq")
```

## Third interaction term: total_nominee_gg * total_academy_nominations
Insignificant
```{r}
# Now I am interested to test if the addition of interaction term is important
final.model <- glm(formula = best_picture_winner ~ popcornmeter + total_academy_award_exclude_best_pic + 
                   nominee_film_edit + best_director_award_before + best_director_award_now + 
                   total_won_gg + budget_norm, family = binomial, data = dat)

# Adding interaction term
interact.model.3 <- glm(formula = best_picture_winner ~ popcornmeter + total_academy_award_exclude_best_pic + 
                   nominee_film_edit + best_director_award_before + best_director_award_now + 
                   total_won_gg + budget_norm + total_nominee_gg * total_academy_nominations, family = binomial, data = dat)

# Anova
anova(final.model, interact.model.3, test = "Chisq")
```

## Forth interaction term: year_2009 with genres
None of the genres*year terms are significant
```{r}
# Adding interaction term
interact.model.4 <- glm(formula = best_picture_winner ~ popcornmeter + total_academy_award_exclude_best_pic + 
                   nominee_film_edit + best_director_award_before + best_director_award_now + 
                   total_won_gg + budget_norm + genre_Drama * post_2009 + genre_HumanInterest * post_2009 +
                     genre_Suspense * post_2009 + genre_Entertainment * post_2009, family = binomial, data = dat)

anova(final.model, interact.model.4, test = "Chisq")
```

## Fifth interaction term: year_2009 with budget_norm
None of the genres*year terms are significant
```{r}
# Adding interaction term
interact.model.5 <- glm(formula = best_picture_winner ~ popcornmeter + total_academy_award_exclude_best_pic + 
                   nominee_film_edit + best_director_award_before + best_director_award_now + 
                   total_won_gg + budget_norm + post_2009*budget_norm, family = binomial, data = dat)

anova(final.model, interact.model.5, test = "Chisq")
```


We hypothesized that the influence of precursor awards and financial performance might be moderated by variables such as genre and budget class (e.g., 'Indie Success' narratives). We tested these interaction effects using Likelihood Ratio Tests. The results were statistically insignificant across all proposed interactions, suggesting that these factors operate additively rather than multiplicatively in predicting Best Picture winners. Consequently, the additive model was retained for its superior parsimony and predictive stability.


We do some robustness check for the final model

```{r}
final.model <- glm(formula = best_picture_winner ~ popcornmeter + total_academy_award_exclude_best_pic + 
                   nominee_film_edit + best_director_award_before + best_director_award_now + 
                   total_won_gg + post_2009*budget_norm, family = binomial, data = dat)
vif(final.model)
summary(final.model)
```


```{r}
residualPlots(final.model)
```


\newpage

Now, we can focus on making the model to have good predictive accuracy.
Our final model has the variable of:

1. popcornmeter
2. total_academy_award_exclude_best_pic
3. nominee_film_edit
4. best_director_award_before
5. best_director_award_now
6. total_won_gg
7. budget_norm

```{r}
# Calculate Odds Ratios and 95% Confidence Intervals
# The percentage change in the winning probability for each unit increase in predictor
or_results <- exp(cbind(OR = coef(final.model), confint(final.model)))
print(or_results)
```


```{r}
library(margins)
me <- margins(final.model)
summary(me)
```


## Visualizing which predictors are important

The Red Line (Value = 1): This is the null hypothesis. If a coefficient’s bar crosses this line, it is not statistically significant at the p < 0.05 level.

The Bars (Confidence Intervals): If the bar touches the red line, the effect is "statistically indistinguishable from zero."

```{r}
summary(final.model)
library(sjPlot)

# This creates a Forest Plot of your final model
plot_model(final.model, 
           show.values = TRUE, 
           value.offset = .3, 
           sort.est = TRUE, 
           title = "Predictors of Best Picture Success",
           axis.title = "Odds Ratio",
           vline.color = "red")
```
## Show that our model has predictive power

```{r}
library(pROC)

# 1. Get predicted probabilities
probabilities <- predict(final.model, type = "response")

# 2. Create the ROC object
roc_obj <- roc(dat$best_picture_winner, probabilities)

# 3. Plot the curve
plot(roc_obj, main = paste("ROC Curve (AUC =", round(auc(roc_obj), 3), ")"))
```

## Likely Overfit :(
```{r}
# 1. Split the data (80% training, 20% testing)
set.seed(123)
train_index <- sample(1:nrow(dat), 0.8 * nrow(dat))
train_data <- dat[train_index, ]
test_data  <- dat[-train_index, ]

# 2. Train the model ONLY on the 80%
test_model <- glm(formula = best_picture_winner ~ popcornmeter + total_academy_award_exclude_best_pic + 
                  nominee_film_edit + best_director_award_before + best_director_award_now + 
                  total_won_gg + budget_norm*post_2009, 
                  family = binomial, data = train_data)

# 3. Predict on the 20% (the "unseen" data)
test_probs <- predict(test_model, newdata = test_data, type = "response")

# 4. Calculate AUC on the unseen data
library(pROC)
test_roc <- roc(test_data$best_picture_winner, test_probs)
print(paste("Validation AUC:", round(auc(test_roc), 3)))

```



## Check the confusion matrix
```{r}
# Use a probability threshold of 0.5 (or lower if your data is very imbalanced)
predicted_class <- ifelse(probabilities > 0.5, 1, 0)
table(Predicted = predicted_class, Actual = dat$best_picture_winner)
# Not using 0.5 due to class imbalance (losing has much more than winning)
```


```{r}
# 1. Find the optimal threshold
# "best" uses Youden's Index (sensitivity + specificity - 1) by default
optimal_result <- coords(roc_obj, "best", ret = "threshold")
optimal_threshold <- optimal_result$threshold

# Print the value so you can see it
print(paste("The optimal threshold is:", round(optimal_threshold, 4)))

# 2. Apply this new threshold to create a more balanced confusion matrix
new_predicted_class <- ifelse(probabilities > optimal_threshold, 1, 0)

# 3. View the new confusion matrix
table(Actual = dat$best_picture_winner, Predicted = new_predicted_class)
```
If the predicted probability is >0.1494, will be classified as winning film.


## Building random forest to explore non-linear relationship
Challenge: class imbalance problem, 175 Losers and 26 Winners:

We check the order to assign the correct weightage
```{r}
levels(as.factor(dat$best_picture_winner))
```

```{r}
library(randomForest)

# Build the forest
# We set ntree to 500 to stabilize the results
# importance = TRUE allows us to extract the importance metrics

# 1. Define weights: 
# The number of losers (0s) vs winners (1s) is roughly 175 to 26.
# We give the '1' class a weight of ~6.7 to balance the importance (175/26 = 6.7)
rf.balanced <- randomForest(as.factor(best_picture_winner) ~ ., 
                            data = dat, 
                            importance = TRUE, 
                            ntree = 500, 
                            mtry = 3,
                            classwt = c(1, 6.73)) # Adjust based on your actual counts

# 2. Check the new confusion matrix
print(rf.balanced)

# 3. Re-generate Variable Importance
varImpPlot(rf.balanced, type = 2)
```

### Examine the robustness of Random Forest
Using rf and see the convergence/stability

+ If the error rate plot (the black line) flattens out before you hit 500 trees, your model has converged. This confirms you didn't overfit by using too few trees.
+ The "OOB estimate of error rate" is a legitimate estimate of your model's accuracy on unseen data. Report this in your findings.

```{r}
# Simply print the model object
print(rf.balanced)

# Plot the error rate over the number of trees
# Look for where the error rate "flattens out"
plot(rf.balanced)
```
Even though we account for balanced class measures, there is still a high error rate for identifying the winner - 92.3%. 
Hence we decided to train the model using the same size of winning and losing datapoints.

```{r}
# Forcing the model to take exactly 26 '0's and 26 '1's for every tree
# This creates a perfectly balanced view for every single decision tree.
rf_strata <- randomForest(as.factor(best_picture_winner) ~ ., 
                          data = dat, 
                          importance = TRUE, 
                          ntree = 500, 
                          mtry = 3,
                          strata = as.factor(dat$best_picture_winner),
                          sampsize = c(26, 26)) # Adjust to your minority class size

# Check the confusion matrix again
print(rf_strata)
varImpPlot(rf_strata, type = 2)
```

# We tried using LASSO onto the logistics regression, but it shrinks everything down to only left with total_academy_award_exclude_best_pic. 

```{r}
# 1. Load the library
library(glmnet)

# 2. Prepare the data
# We remove the response variable and 'rating' as you did in the full model
# model.matrix handles categorical variables by creating dummy variables automatically
x <- model.matrix(best_picture_winner ~ . - rating, data = dat)[, -1] # [, -1] removes the intercept column
y <- dat$best_picture_winner

# 3. Use Cross-Validation to find the optimal Lambda (penalty strength)
# family = "binomial" specifies logistic regression
cv_lasso <- cv.glmnet(x, y, family = "binomial", alpha = 1) 

# 4. Plot the cross-validation results to see the error vs. Log(Lambda)
plot(cv_lasso)

# 5. Extract the "Best" Lambda
# lambda.min: value that gives minimum mean cross-validated error
# lambda.1se: most regularized model such that error is within 1 std err of the minimum
best_lambda <- cv_lasso$lambda.min 

# 6. View the selected variables (coefficients)
# Variables with a "." are the ones LASSO has "zeroed out" (deselected)
coef(cv_lasso, s = best_lambda)
```
```{r}
# Use the formula syntax inside model.matrix to generate the interaction columns
# This creates columns for: post_2009, budget_norm, and post_2009:budget_norm
x_interact <- model.matrix(best_picture_winner ~ . - rating + post_2009:budget_norm, data = dat)[, -1]
y <- dat$best_picture_winner

# Run CV-Lasso as usual
cv_lasso_interact <- cv.glmnet(x_interact, y, family = "binomial", alpha = 1)

# Check which coefficients (including the interaction) survived
coef(cv_lasso_interact, s = "lambda.min")

```
```{r}
# 1. Identify variables selected by LASSO (e.g., at lambda.min)
# Let's say LASSO selected: budget_norm, nominee_film_edit, and total_won_gg

# 2. Run a standard GLM using ONLY those variables
lasso_vars_model <- glm(best_picture_winner ~ budget_norm + nominee_film_edit + total_won_gg, 
                        data = dat, family = binomial)

# 3. Check the AIC
AIC(lasso_vars_model)

```



